Parsing Postfix log files

### 2009-03-29 doc/logparser-results.tex
Corrections after proofreading.

@@

XXX COMPARE PARSING + ACTIONS TO JUST PARSING\@.

@@

XXX COMPARE BEST AND WORST POSSIBLE RESULTS WITH REAL RESULTS\@.

@@

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} shows that as expected there are far more
mails received on weekdays than at weekends.  The mail server in question
is a production mail server handling mail for a university department; the
benefit of using this server is that the log files from it exhibit the
idiosyncrasies and peculiarities a mail server in the wild must deal with,
but the downside is that significantly altering the configuration just to
log messages from a different Postfix component is not an option.  Median
log file size is 50MB, containing 285,000 log lines --- large scale mail
servers would have much larger log files.  Note that \graphref{Mails
received per day} and \tableref{Number of mails received per day:
statistics} show the number of mails received by \acronym{SMTP} only; in
particular the mail loops noticeable in other graphs do not contribute to
these figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times and the parsing times averaged.  Saving results to the
database was disabled for the test runs, as that dominates the run time of
the program, and the tests are aimed at measuring the speed of the parser
rather than the speed of the database and the disks the database is stored
on.  The computer used for test runs was a Dell Optiplex 745, equipped
with:

\begin{tabular}[]{ll}

    CPU         & One dual core 2.40GHz Intel\textregistered{}
                    Core\texttrademark{}2 CPU,                      \\
                & with 32KB L1 cache and 4MB L2 cache.              \\
    RAM         & 2GB 667 MHz DDR RAM\@.                            \\
    Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\

\end{tabular}

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and tens of thousands of log
lines for each of the \numberOFlogFILES{} log files.  All three lines run
roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor} which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, providing empirical evidence that the algorithm
scales linearly.  The ratio increases (i.e.\ improves) for log files 22 and
62--68 despite their larger than average size (shown in \graphref{parsing
time vs file size vs number of log lines graph}); this is explained fully
in \sectionref{Explaining the peaks in log file size}.

@@

\Graphref{parsing time vs file size vs number of log lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline{}
\tab{}\texttt{$\backslash$username, username@domain} \newline{} This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this instance is the address the mail was originally sent
to, creating an infinite loop.  To prevent this happening Postfix examines
the Delivered-To header in the mail, and if the mail has already been
delivered to the current address it is bounced back to the sender with the
error message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Ordinarily this works well, but unfortunately
in this instance the user noticed they had not received any mail in a while
and opted to send a test mail to themselves, causing a loop not caught by
Postfix:

XXX REWRITE THIS SECTION SO THAT IT IS CORRECT\@: POINT 4 SAYS THAT THE
ENVELOPE SENDER CHANGES, IS THAT CORRECT\@?

    \item Postfix accepts a mail from username@domain, for username@domain.

    \item Postfix delivers the mail to the local mailbox and
        username@domain, as instructed by the user's forwarding
        instructions.  The forwarded mail has a
        \texttt{Delivered-To:~username@domain} header added, and the
        envelope sender address is username@domain.

    \item Postfix accepts the mail for username@domain, but during delivery
        it notices that the \texttt{Delivered-To} header already contains
        the address it is currently delivering to, and therefore sends a
        bounce notification with sender address \texttt{<>}\glsadd{<>} to
        the original sender: username@domain.

    \item Postfix accepts the bounce notification and delivers it to both
        the local mailbox and to username@domain, as instructed by the
        user's forwarding instructions.  A
        \texttt{Delivered-To:~username@domain} header is added to the
        forwarded bounce notification, XXX which now has an envelope sender
        address of username@domain.   XXX IS THIS CORRECT\@?

    \item Postfix accepts the forwarded bounce notification but during mail
        delivery it notices that the \texttt{Delivered-To} header already
        contains the address currently being delivered to, and sends a
        bounce notification to the sender: username@domain.

    \item Repeat from step two; this will continue indefinitely unless an
        administrator intervenes and deletes the appropriate mails from the
        queue.

The sequence described above occurs extremely rapidly because Postfix does
not have to deliver the mail to an external system, so mails are delivered,
bounced, and generated as fast as the disks can keep up, resulting in a
huge volume of log lines.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% recognise \input{build/include-top-ten-hits}\% of the log lines, with
the remaining log lines split across the other 90\% of the rules (as shown
in \graphref{rule hits graph}).  Assuming that the distribution of log
lines is reasonably consistent over time, parser efficiency should benefit
from trying more frequently recognising rules before those that recognise
less frequently.  To test this hypothesis three full test runs were
performed with different rule orderings:

@@

    \item [Shuffle] This is intended to represent a randomly ordered rule
        set.  The rules will be shuffled once before use and will retain
        that ordering until the parser exits.  Note that the ordering will
        change every time the parser is executed, so 10 different rule
        orderings will be generated for each log file in the test run.

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general the
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of log lines versus parsing time in
\graphref{parsing time vs file size vs number of log lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files used
when generating the results in this thesis is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because the Postfix parser is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses logs generated
by later Postfix versions.  A second set of test runs was performed using
the minimum ruleset, and the parsing times compared to those generated
using the full ruleset: the percentage parsing time increase when using the
full ruleset instead of the minimal ruleset for optimal, shuffled and
reverse orderings is shown in \graphref{Percentage parsing time increase of
maximum ruleset over minimum ruleset}.

@@

Clearly the increased number of rules has a noticeable performance impact
with reverse ordering, and a lesser impact with shuffled ordering.  The
optimal ordering (where the rules that recognise log lines most frequently
are tried first) shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  Log
files 22 and 62--68 show much smaller increases for shuffled and reversed
ordering than other log files do, because the majority of log lines in
those log files are parsed by Postfix components with a small number of
parsing rules, so removing unnecessary rules has little effect on the total
number of rules used.  These results show that the architecture scales very
well as the number of rules increases, and that optimally sorting the rules
is an important optimisation contributing to this scalability.

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and
optimised to improve the speed of matching.  This compilation and
optimisation takes CPU time ---in many cases it takes far more CPU time
than the actual matching.  If the interpreter is certain that a regex
will not change it will automatically cache the compilation results for
later use.   The results of compiling a dynamically generated regex can
be cached and used in preference to the original regex, but it is the
responsibility of the programmer to do this.  The Postfix parser loads the
rules it uses from a database, and thus each rule's regex is compiled
and the result cached to reduce parsing time.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the impact that not caching compiled regexes has on parser
performance: on typical log files the parsing time when not caching
compiled regexes is 400--600\% of the parsing time when caching.
Caching compiled regexes is probably the single most effective
optimisation possible in the parser's implementation, and was quite simple
to implement.

@@

Two large dips can be seen in \graphref{Increase in parsing time when not
caching compiled regexes graph} at log files 22 and 62--68, corresponding
to the spikes in log file size in \graphref{parsing time vs file size vs
number of log lines graph}.  The distribution of log lines across rules
when there is a mail loop is much different to the norm, and the average
number of rules consulted per log line is much lower: this results in far
fewer regex compilations per line when there is a mail loop, and a
corresponding decrease in the average parsing time for a single log line.
The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for different
groups of log files are summarised in \tableref{Increase in parsing time
when not caching compiled regexes table}.

\showtable{build/include-cached-regexes-vs-discarded-regexes} {Increase in
parsing time when not caching compiled regexes} {Increase in parsing time
when not caching compiled regexes table}

@@

>>>>

% XXX IS CHANGING FIGURE TO GRAPH A GOOD IDEA?
\renewcommand{\figurename}{Graph}

@@

@@

@@

XXX THINK ABOUT THE ORDERING OF TABLES AND GRAPHS\@; MAYBE REARRANGE THEM
SO THAT THEY APPEAR IN THE ORDER THEY ARE REFERENCED\@?

XXX SHOULD I USE $\backslash{}$clearpage{} OR SOMETHING TO AFFECT
PLACEMENT\@?  GRAPHS AND TABLES TEND TO MOVE QUITE FAR IN SOME OF THE
SUBSECTIONS, THOUGH MAKING THE TEXT LONGER MIGHT HELP TOO\@.  IDEALLY I
WANT TO HAVE TEXT, THEN ALL GRAPHS AND TABLES, THEN THE NEXT SECTION\@; I
DO NOT LIKE THAT GRAPHS/TABLES SOMETIMES APPEAR BEFORE THEY ARE
REFERENCED\@.  I MIGHT NEED TO REDUCE THE HEIGHT OF THE GRAPHS TOO\@.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} shows that, as expected, far more mails
are received on weekdays than at weekends.  The mail server in question is
a production mail server handling mail for a university department; the
benefit of using this server is that its log files exhibit the
idiosyncrasies and peculiarities a mail server in the wild must deal with,
but the downside is that significantly altering the configuration to
accommodate this project is not an option.  Median log file size is 50MB,
containing 285,000 log lines; large scale mail servers would have much
larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \acronym{SMTP} only; in particular the mail loops
noticeable in later graphs do not contribute to these figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, and the parsing times from the 10 runs averaged.  Saving
results to the database was disabled for the test runs, because that
dominates the run time of the program, and the tests are aimed at measuring
the speed of \parsername{} rather than the speed of the database and the
disks the database is stored on.  The computer used for test runs was a
Dell Optiplex 745, shown in \tableref{Computer used to generate
statistics}.

\begin{table}[htbp]
    \caption{Computer used to generate statistics}
    \empty{}\label{Computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component in use                                   \\
        \tablemiddleline{}%
        CPU         & One dual core 2.40GHz Intel\textregistered{}
                        Core\texttrademark{}2 CPU,                      \\
                    & with 32KB L1 cache and 4MB L2 cache.              \\
        RAM         & 2GB 667 MHz DDR RAM\@.                            \\
        Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly.  The
ratios increase (i.e.\ improve) for log files 22 and 62--68 despite their
larger than average size (shown in \graphref{parsing time vs file size vs
number of log lines graph}); this is explained fully in
\sectionref{Explaining the peaks in log file size}.

@@

XXX SHOULD THIS BE A SEPARATE SECTION OR INCLUDED IN THE PREVIOUS ONE\@?
MAYBE INCLUDE GRAPHS OF RULES PER FILE OR SOMETHING TO EXTEND THIS\@?

XXX REWRITE WHEN STATS RUNS HAVE FINISHED\@; REWRITE ALL THE REFERENCES
TOO\@.

\Graphref{parsing time vs file size vs number of log lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline{}
\tab{}\texttt{$\backslash$username, username@domain} \newline{} This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this instance is the address the mail was originally sent
to, creating an infinite loop.  To prevent this happening Postfix examines
the Delivered-To header in the mail, and if the mail has already been
delivered to the current address it is bounced back to the sender with the
error message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Normally this works well, but the forwarding
method in use at that time stripped the Delivered-To header from the mail,
breaking loop detection.

When the recipient in a mail loop is local, Postfix does not have to
deliver the mail to an external system, so mails are delivered as fast as
the disks can keep up, resulting in a huge volume of log lines.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% recognise \input{build/include-top-ten-hits}\% of the log lines, with
the remaining log lines split across the other 90\% of the rules (as shown
in \graphref{rule hits graph}).  Assuming that the distribution of log
lines is reasonably consistent over time, \parsernames{} efficiency should
benefit from trying more frequently recognising rules before those that
recognise less frequently.  To test this hypothesis three full test runs
were performed with different rule orderings:

@@

    \item [Shuffle] This ordering is intended to represent a randomly
        ordered rule set.  The rules will be shuffled once before use and
        will retain that ordering until the parser exits.  Note that the
        ordering will change every time the parser is executed, so 10
        different rule orderings will be generated for each log file in the
        test run.

@@

\newpage{} % XXX VARIOREF PROBLEM

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of log lines vs.\ parsing time in
\graphref{parsing time vs file size vs number of log lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full \parsername{} ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files used
when generating the results in this thesis is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because \parsername{} is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses logs generated
by later Postfix versions.  A second set of test runs was performed using
the minimum ruleset, and the parsing times compared to those generated
using the full ruleset: the percentage parsing time increase when using the
full ruleset instead of the minimal ruleset for optimal, shuffled and
reverse orderings is shown in \graphref{Percentage parsing time increase of
maximum ruleset over minimum ruleset}.

@@

Clearly the increased number of rules has a noticeable performance impact
with reverse ordering, and a lesser impact with shuffled ordering.  The
optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
XXXXXXXXXXXX IMPROVE THIS EXPLANATION\@.  Log files 22 and 62--68 show much
smaller increases for shuffled and reversed ordering than other log files
do, because the majority of log lines in those log files are parsed by
Postfix components with a small number of parsing rules, so removing
unnecessary rules has little effect on the total number of rules used.
These results show that the architecture scales very well as the number of
rules increases, and that optimally sorting the rules is an important
optimisation contributing to this scalability.

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and optimised
to improve the speed of matching.  This compilation and optimisation takes
CPU time: in many cases it takes far more CPU time than the actual
matching.  If the interpreter is certain that a regex will not change, it
will automatically cache the compilation results for later use.   The
results of compiling a dynamically generated regex can be cached and used
in preference to the original regex, but it is the responsibility of the
programmer to do this.  \parsername{} loads the rules it uses from a
database, so it compiles each rule's regex and caches the result to reduce
parsing time.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the impact that not caching compiled regexes has on parser
performance: on typical log files, the parsing time when not caching
compiled regexes is 400--600\% of the parsing time when caching.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement.

@@

XXX UPDATE THIS\@.  Two large dips can be seen in \graphref{Increase in
parsing time when not caching compiled regexes graph} at log files 22 and
62--68, corresponding to the spikes in log file size in \graphref{parsing
time vs file size vs number of log lines graph}.  The distribution of log
lines across rules when there is a mail loop is much different to the norm,
and the average number of rules consulted per log line is much lower: this
results in far fewer regex compilations per line when there is a mail loop,
and a corresponding decrease in the average parsing time for a single log
line.  The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for different
groups of log files are summarised in \tableref{Increase in parsing time
when not caching compiled regexes table}.

\showtable{build/include-cached-regexes-vs-discarded-regexes}{Increase in
parsing time when not caching compiled regexes}{Increase in parsing time
when not caching compiled regexes table}

@@

XXX REWRITE WHEN THIS CHAPTER IS FINISHED\@.
